## Exercises


::: {#exr-}
How can you tell if an object is a tibble?
:::



::: {#exr-}
Compare and contrast the following operations on a `data.frame` and equivalent `tibble`. What is different? Why might the default data frame behaviors cause you frustration?

```{r}
#| eval: false
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

```


:::



::: {#exr-}
If you have the name of a variable stored in an object, e.g., `var <- "xyz"`, how can you extract the reference variable from a tibble? You may use the following codes to get a tibble.

```{r}
tbl <- tibble(abc = 1, xyz = "a")
```

:::


::: {#exr-}
Practice referring to nonsyntactic names in the following data.frame by:

a. Extracting the variable called 1.
b. Creating a new column called 3, which is 2 divided by 1.
c. Renaming the columns to one, two, and three:

```{r}
annoying <- tibble(
`1` = 1:10,
`2` = `1` * 2 + rnorm(length(`1`))
)
```
:::




::: {#exr-}
Both `unite()` and `separate()` have a `remove` argument. What does it do? Why would you set it to `FALSE`?

:::



::: {#exr-}
Use `flights` dataset. Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
:::